\Large{Learning to Pivot with Adversarial Networks}

### 2016-11-02 tex/paper.tex
last changes?

We begin with a family of data generation processes $p(X,Y,Z)$, where $X \in
\mathcal{X}$ are the data, $Y\in \mathcal{Y}$ are the target labels, and $Z\in
\mathcal{Z}$ are the nuisance parameters that can be continuous or categorical.
We assume training data $\{x_i, y_i, z_i\}_{i=1}^N$. Let us assume that prior to incorporating the effect of uncertainty
in $Z$, our goal is to learn a regression function $f : \mathcal{X} \mapsto
{\cal S}$ of parameters $\theta_f$ (e.g., a neural network-based probabilistic
classifier) that minimizes a loss ${\cal L}_f(\theta_f)$ (e.g., the
cross-entropy).

@@

As stated in Eqn.~\ref{eqn:criterion}, the pivotal quantity criterion is
imposed with respect to the $p(X|Z)$ with $Y$ is marginalized out. In some situations
however (see e.g., Sec.~\ref{sec:hep}), class conditional independence of $f(X;
\theta_f)$ on the nuisance $Z$ is preferred, which can then be stated as
requiring
\begin{equation}\label{eqn:criterion-class}
    p(f(X ; \theta_f) = s | z, y ) = p(f(X ; \theta_f) = s | z^\prime, y )
\end{equation}
for one or several specified values $y \in {\cal Y}$.

@@

Proposition~\ref{prop:2} suggests that if at each step of
Algorithm~\ref{alg:adversarial-training} the adversary $r$ is allowed to reach
its optimum given $f$ (e.g., by setting $K$ sufficiently high) and if $f$ is
updated to improve ${\cal L}_f(\theta_f) -  H({Z|f(X;\theta_f)})$ with
sufficiently small steps, then $f$ should converge to a classifier that is both
optimal and pivotal, provided such a classifier exists. Let us note
that despite the former theoretical characterization of the minimax solution
of Eqn.~\ref{eqn:min_thetaf}, formal guarantees of
convergence towards that solution by
Algorithm~\ref{alg:adversarial-training} in the case where a finite number $K$
of steps is taken for $r$ remains to be proven.

@@

In machine learning, learning a pivotal quantity can be related to the problem
of domain adaptation
\citep{blitzer2006domain,pan2011domain,gopalan2011domain,gong2013connecting,baktashmotlagh2013unsupervised,ajakan2014domain,ganin2014unsupervised},
where the goal is often stated as trying to learn a domain-invariant
representation of the data. Likewise, our method also relates to the problem of
enforcing fairness in classification
\citep{zemel2013learning,feldman2015certifying,EdwardsS15,zafar2015fairness},
which is stated as learning a classifier that is independent of some chosen
attribute such as gender, color or age. For both families of methods, the
problem can equivalently be stated as learning a classifier which is a pivotal
quantity with respect to either the domain or the selected feature. In this
context, \citep{ganin2014unsupervised,EdwardsS15} are certainly among the
closest to our work, in which domain invariance and fairness are enforced
through an adversarial minimax setup composed of a classifier and an adversary
discriminator. Following this line of work and without making assumptions on the
prior $p(Z)$, our method can be regarded as a generalization that also supports
a continuously parametrized family of domains or as enforcing fairness over
continuous attributes.

>>>>

We begin with a family of data generation processes $p(X,Y,Z)$, where $X \in
\mathcal{X}$ are the data, $Y\in \mathcal{Y}$ are the target labels, and $Z\in
\mathcal{Z}$ are the nuisance parameters that can be continuous or categorical.
% We assume training data $\{x_i, y_i, z_i\}_{i=1}^N$.
Let us assume that prior to incorporating the effect of uncertainty
in $Z$, our goal is to learn a regression function $f : \mathcal{X} \mapsto
{\cal S}$ of parameters $\theta_f$ (e.g., a neural network-based probabilistic
classifier) that minimizes a loss ${\cal L}_f(\theta_f)$ (e.g., the
cross-entropy).

@@

As stated in Eqn.~\ref{eqn:criterion}, the pivotal quantity criterion is
imposed with respect to $p(X|Z)$ where $Y$ is marginalized out. In some situations
however (see e.g., Sec.~\ref{sec:hep}), class conditional independence of $f(X;
\theta_f)$ on the nuisance $Z$ is preferred, which can then be stated as
requiring
\begin{equation}\label{eqn:criterion-class}
    p(f(X ; \theta_f) = s | z, y ) = p(f(X ; \theta_f) = s | z^\prime, y )
\end{equation}
for one or several specified values $y \in {\cal Y}$.

@@

Proposition~\ref{prop:2} suggests that if at each step of
Algorithm~\ref{alg:adversarial-training} the adversary $r$ is allowed to reach
its optimum given $f$ (e.g., by setting $K$ sufficiently high) and if $f$ is
updated to improve ${\cal L}_f(\theta_f) -  H({Z|f(X;\theta_f)})$ with
sufficiently small steps, then $f$ should eventually
converge to a classifier that is both
optimal and pivotal. Provided this classifier exists, the adversarial term ${\cal L}_r$
can therefore be regarded as a way to select among
the class of all optimal classifiers a function $f$ that is also pivotal.
Despite the former theoretical characterization of the minimax solution
of Eqn.~\ref{eqn:min_thetaf}, let us note that formal guarantees of
convergence towards that solution by
Algorithm~\ref{alg:adversarial-training} in the case where a finite number $K$
of steps is taken for $r$ remains to be proven.

@@

In machine learning, learning a pivotal quantity can be related to the problem
of domain adaptation
\citep{blitzer2006domain,pan2011domain,gopalan2011domain,gong2013connecting,baktashmotlagh2013unsupervised,ajakan2014domain,ganin2014unsupervised},
where the goal is often stated as trying to learn a domain-invariant
representation of the data. Likewise, our method also relates to the problem of
enforcing fairness in classification
\citep{zemel2013learning,feldman2015certifying,EdwardsS15,zafar2015fairness},
which is stated as learning a classifier that is independent of some chosen
attribute such as gender, color or age. For both families of methods, the
problem can equivalently be stated as learning a classifier which is a pivotal
quantity with respect to either the domain or the selected feature. As an
example, unsupervised domain adaptation with labeled data from a source domain
and unlabeled data from a target domain can be recast as learning a predictive
model $f$ (i.e., trained to minimize ${\cal L}_f$ evaluated on labeled source
data only) that is also a pivot with respect to the domain $Z$ (i.e., trained to
maximize ${\cal L}_r$ evaluated on both source and target data). In this
context, \citep{ganin2014unsupervised,EdwardsS15} are certainly among the
closest to our work, in which domain invariance and fairness are enforced
through an adversarial minimax setup composed of a classifier and an adversary
discriminator. Following this line of work and without making assumptions on the
prior $p(Z)$, our method can be regarded as a generalization that also supports
a continuously parametrized family of domains or as enforcing fairness over
continuous attributes.
